Procedural generation of levels for the angry birds videogame using evolutionary computation\\

### 2019-05-15 IEEE_ConferenceOnGames/article_english.tex
soa

% What is the problem?
Procedural content generation provides a peculiar case of study because of the way
% Is there a reference? - MG
content has to be made for a each particular type of game. In this use case 
we have those restrictions dictated by the game, and also 
those imposed by a competition track \cite{Renz}.% Reference the paper of the competition  have provide an interesting 
The challenge for the participating algorithms is to be flexible on the  
pieces that can be used and the runing time of the algorithm, as they are determined at runtime. 
%their interactions with one another because of 
%gravity and the possible effects the trajectory of a shot from a player can affect 
%all of this.
% How we solve it? (summary)   
In order to tackle this case study, we proposed the use a genetic algorithm capable 
of creating game levels having the objective of being both \textit{interesting} and 
\textit{playable}.

@@

% Not important - MG
%As mentioned above in this paper we use procedural content generation (PCG) 
%which denotes the way to automatically generate
%content by algorithms instead of generating the same content by manual means
%such as with different personnel.

%In the videogame industry many developers use a PCG style software that produces
%raw images and designs used by other personnel to improve it and use it in order
%to reduce the development cycles, and the productions cost. 
%PCG as various focus areas that resolve around a video game, however in this paper 
%the main focus is the \textit{Level design} area which as the objective to design 
%the different areas of a game in which a player can interact with, since in the 
%angry birds video game a level is represented by a scene then the system will 
%have to create the entities (birds and pigs) and objects (pieces) in it in order 
%to create a playable level.

We propose a basic Open-Ended Evolution (OEE) algorithm, a variant of genetic evolution algorithm where
the objective differs from that of a regular genetic algorithm,  in this case the
final objective is not to reach a final state of the evolution (an optimal solution) but to continue
evolving and creating new types of entities, perhaps more complex on each new generation.
Solutuion entities have the possibility to evolve not only inside the frame of
their type, but  they can create entirely new groups \cite{Standish2003}. T. Taylor et al. \cite{Taylor2016,Taylor}
defines OEE as a system capable of creating novelty beyond a point where
nothing changes on the current group of interest rather than just converging to
a stable or quasi-stable state.

For this project, OEE is defined simply as a genetic evolution process with
different clasess of objects, with compositional and taxonomic relationships between them.
Objects of each class can evolve in parallel or as in this case, independenly of each other.   
The idea is based on the object oriented model, in which complex behavior can be model 
using these principles.

% I tried to mix the two sections together while expanding the first part of 
% the introduction JS

%In Section 2, a background of the subjects used on this project will be
%presented, this subjects, while not all being used on by different authors on
%their works provide a great help on the development of the proposed system,
% Jaime, remember there is no Background need it in this type of paper 
% Maybe that will be the introduction - MG

@@

\section{Overview of existing work}

In this Section we present some of the different approaches to the procedural
generation of content for the angry birds video game.

M. Kaidan et al. \cite{Kaidan2015} propose a model in which the generation of
content is controlled by the skill of a player while actively playing the game,
the moment the player finishes a certain level the algorithm obtains the score
and generates the next level to be played.

J. Renz et al. \cite{Renz} created an article in which the angry birds level
generation contest mechanics of the IEEE Conference on Computational
Intelligence and Games are based on, it explains some of the solving methods
proposed by previous competitors.

Y. Jiang et al. \cite{Jiang2017} proposed the use of predefined letter style
patterns and combined to create a small pool of words and letters used to be combined in
order to create levels with text on them, the participants would play the game
and after that they would be asked if they liked the layout of the levels.

>>>>

% What is the problem?
Procedural content generation provides a peculiar case of study because of the
way
% Is there a reference? - MG
content has to be made for a each particular type of game. In this use case we
have those restrictions dictated by the game, and also those imposed by a
competition track \cite{Renz}.% Reference the paper of the competition  have
provide an interesting The challenge for the participating algorithms is to be
flexible on the pieces that can be used and the runing time of the algorithm, as
they are determined at runtime. 
%their interactions with one another because of gravity and the possible effects
%the trajectory of a shot from a player can affect all of this. How we solve it?
%(summary)   
In order to tackle this case study, we proposed the use a genetic algorithm
capable of creating game levels having the objective of being both
\textit{interesting} and \textit{playable}.

@@

We propose a basic Open-Ended Evolution (OEE) algorithm, a variant of genetic
evolution algorithm where the objective differs from that of a regular genetic
algorithm,  in this case the final objective is not to reach a final state of
the evolution (an optimal solution) but to continue evolving and creating new
types of entities, perhaps more complex on each new generation. Solutuion
entities have the possibility to evolve not only inside the frame of their type,
but  they can create entirely new groups \cite{Standish2003}. Taylor et al.
\cite{Taylor2016,Taylor} defines OEE as a system capable of creating novelty
beyond a point where nothing changes on the current group of interest rather
than just converging to a stable or quasi-stable state.For this project, OEE is
defined simply as a genetic evolution process with different clasess of objects,
with compositional and taxonomic relationships between them. Objects of each
class can evolve in parallel or as in this case, independenly of each other. The
idea is based on the object oriented model, in which complex behavior can be
model using these principles. The rest of the paper is as follows.

%% TO DO: This needs to be adjusted to the final version. We need to use
% \ref for the numbers - MG

@@

\section{State of the art}
\label{SoA}
In this Section we present PCG works relevant to the Level generation for the
Angry Birds video game. Renz et al. \cite{Renz, Renz2015TheAB} organized an AI
competition with the aim of developing intelligent agents that could play new
Angry Birds levels better than the best human players. The competition was then
collocated in other AI conferences. In order to train and test these agents,
more novel and some times difficult levels must be generated. A related
competition was then proposed by Stephenson et al. \cite{Stephenson} this time
with the goal of developing a level generator for the game. Several generators
have been proposed in the following years of competition. M. Kaidan et al.
\cite{Kaidan2015} propose a model in which the generation of content is
controlled by the skill of a player while actively playing the game, the moment
the player finishes a certain level the algorithm obtains the score and
generates the next level to be played. Y. Jiang et al. \cite{Jiang2017} proposed
the use of predefined letter style patterns and combined to create a small pool
of words and letters used to be combined in order to create levels with text on
them, the participants would play the game and after that they would be asked if
they liked the layout of the levels. A search based approach has been also used
in several works, recently a
